Type Error Telemetry at Scale

### 2023-09-29 programming23/paper.tex
sec 4 getting started

Roblox Studio collected type error telemetry in Spring 2023,
between February and April.
Every Studio session had a small random chance of generating
telemetry~(\cref{s:telemetry-records}).
The chosen sessions generated a record whenever the user switched modules and
randomly on each keystroke.
In total, we collected 1.5 million telemetry records.
Roughly two-thirds of records are due to keystrokes.

\Cref{f:records-per-hour} provides a time-ordered distribution of the data.
Each vertical bar counts the number of records generated per hour according to
a client-side timestamp.\todo{Do we need all this detail?}
The bars are labeled with a California time zone because that is the locale of
the Roblox servers, which parsed the timestamps.
Most buckets contain roughly one thousand records, but some have very few
and a handful of others exceed three thousand.
%% \FILL{} investigate one tower, what's happening?? all one session?
There is a sharp drop in early April when the experiment was finished
and the telemetry probe was removed (though there is a long tail as
developers uopdated Roblox Studio).
We do not know which timezone the records originated in, but since the counts
tend to peak midday California time it seems that the majority
of Roblox developers are following a Western Hemisphere schedule.
The counts for weekends (shaded regions) are often the tallest,
as Roblox has a significant school-aged creator community.

@@

A close inspection of the data revealed a few anomalies.
For the most part, we omit the strange data from futher
analysis.\todo{I'm inclined to rewrite this to be much more concise.}\todo{agreed}

First, some records have odd timestamps according to the server's records.
Every telemetry record comes with two timestamps: a client-side timestamp
from when Roblox Studio created it and a server-side timestamp from when
the server posted it to its database.
The server timestamps are little use for our analysis because they bunch
together; it is common for several records to have identical server
timestamps.
However, server timestamps are useful for auditing because they should
be relatively close in value to the client timestamps.
For a small number of records (<100? \FILL{}), though, the client is off by a large
margin: up to one week behind and up to one day ahead of the server.
We attribute the large delays to issues when sending telemetry (lost internet
connection, etc.) and smaller offsets to incorrect time settings on the client
machine.
In any event, we do not filter these records from further
analysis.\todo{To be clear, you do NOT filter? Because the section is
  about what you DO filter.}

Second, some records have identical (client) timestamps.
This can happen legitimately if type analysis runs several times
in the same millisecond, but it is unlikely that a human would
enter keystrokes or switch modules quickly enough.
We keep the first record with a distinct timestamp and discard the others.
%% TODO wait, better deduplicate by session and not globally!!!

%% out/negative-edit-range.txt
%% actual value is NOT negative, but large positive over 4 billion
%% example: 4294967293
Third, the edit ranges are negative in a few hundred records~(1,533).
This is likely due to the user deleting a large block of code.
Since the issue affects so few records, we omit the negative ranges
from analysis (but, we do use other parts of the records; for example
the timestamps influence~\cref{f:records-per-hour}).

@@

Three important characteristics of the recorded data are the size of the codebases,
the length of the sessions, and the number of type
errors.
We discuss these in turn.
Codebase size and session length follow power-law
distributions,\todo{Please confirm w/ a statistical test. See Clauset
  et al.: arXiv:0706.1062.} with
many small items and a few very large items.
The number of type analysis errors is somewhat low~(explained in the next
section), but sufficient to proceed with futher analysis.

@@

\begin{figure}[t]\centering
  % ;; (list* (max* vv) (median < vv) mm (stddev/mean mm vv) (percentile* vv)))
  % ;; percentile* = 0.95 -- 0.99
  % #hash((editrange . (1156036 926 3007043/817 30975.40553032358 (0.95 8220) (0.96 9858) (0.97 13656) (0.98 18956) (0.99 34725)))
  %       (event-count . (6079 138 38689/135 582.5752201972966 (0.95 960) (0.96 1174) (0.97 1304) (0.98 1836) (0.99 3302)))
  %       (files . (54884 7678 55257996/4639 11853.650098459446 (0.95 40029) (0.96 42771) (0.97 45417) (0.98 48688) (0.99 51761)))
  %       (lines . (1089963 3115 38928947/5992 22364.75846303099 (0.95 18561) (0.96 25892) (0.97 27014) (0.98 29725) (0.99 50547)))
  %       (timespan . (1387897376 845648 591343604248/185705 15724745.722577972 (0.95 10696064) (0.96 12715468) (0.97 15596940) (0.98 21153924) (0.99 35450460))))
  %% NOTE: 3 records have +1M lines, all from one nocheck session, all with 3 files
  %% - session "30,762,216,447,324"
  %% NOTE: 133 records have +52K files, all from one nocheck session
  %% - session "330,172,870,224,081" 
  %% (2nd biggest has 41K files "31,650,473,913,529")

@@

  \caption{Size of analyzed code: number of files, number of lines, and lines in edit range}
  \label{f:codebase-size}
\end{figure}

\Cref{f:codebase-size} summarizes the size of projects in the dataset.
The top half of the figure reports the mean, standard deviation, median,
and \percentile{99} for the number of lines
and lines in edit range.\footnote{The data also contains the number of files in
each project, but these numbers are difficult to relate to typechecking because
projects may contain hundreds of files that define data assets.  For the
record, the median file count is 7,678 and the 99th percentile is 51,761 files.
}
The bottom half shows zoomed-in distributions of the line and edit range
sizes.
For example, the x-axis of the first plot ranges from 0 to 1,000 lines and
the y-axis counts up to 1,000 telemetry records.

There is a huge amount of variation across projects.
The largest ones have over 50,000 lines of code
while the smallest have merely 1 line of code.
Unsurprisingly, the numbers come with large standard
deviations~([std] in~\cref{f:codebase-size}).
The median values are more reasonable, with roughly 3,000 lines of code
and 1,000 lines in edit ranges.

@@

\Cref{f:session-size} outlines the size of the Roblox Studio sessions in the
data.
There are two useful notions of size: real time in seconds (between the first
and last record) and the total number of records.
The top of~\cref{f:session-size} reports the mean, standard deviation,
median, and \percentile{99}; the bottom shows size distributions, zoomed in
to focus on at most 1,000 units (x-axis) and 1,000 sessions
(y-axis).\todo{Again, confused.}

>>>>

Roblox Studio collected type error telemetry between February
and April 2023.
Every Studio session had a small random chance of generating
telemetry~(\cref{s:telemetry-records}).
The chosen sessions generated a record whenever the user switched modules and
randomly on each keystroke.
In total, we collected over 1.5 million telemetry records.
Roughly two thirds of records are due to keystrokes.

\Cref{f:records-per-hour} provides a time-ordered distribution of the data.
Each vertical bar counts the number of records generated per hour according to
a client-side timestamp.
%% \todo{Do we need all this detail?}
%% the vertical bars suggest we have lots of day-job developers, will point this out below
The bars are labeled with a California time zone because that is the locale of
the Roblox servers, which parsed the timestamps.
Most buckets contain roughly one thousand records, but the number
varies between a few dozen at minimum at over three thousand.
%% \FILL{} investigate one tower, what's happening?? all one session?
There is a sharp drop in early April when the experiment was finished
and the telemetry probe was removed, though there is a long tail as
developers updated {Roblox Studio}.
We do not know which timezone the records originated in, but since the counts
tend to peak midday California time it seems that the majority
of Roblox developers are following a Western Hemisphere schedule.
The regular pattern of the peaks furthermore suggests that many creators
follow a regular schedule; this may be due to a large community of
professional developers.
The counts for weekends (shaded regions) are often the tallest,
as Roblox has a significant school-aged creator community.

@@

A close inspection of the data revealed a two anomalies
that we removed from further analysis:
\begin{enumerate}
  \item
    Some records have identical timestamps and session IDs.
    We keep the first such record and discard the rest.
    (A human would have to switch modules twice or enter several keystrokes
     within one millisecond to generate two such records legitimately.)

   \item
     A few hundred records~(1,533) have negative edit ranges.
     These are likely due to large deletions.
     Since the issue affects so few records, we simply ignore
     their edit ranges. We do, however, use their overall error counts and other uncorrupted fields.

\end{enumerate}

%% not important for submission
% One related observation, for which \emph{do not} sanitize the data, is
% that some records have odd timestamps according to the server's records.
% Every telemetry record comes with two timestamps: a client-side timestamp
% from when Roblox Studio created it and a server-side timestamp from when
% the server posted it to its database.
% The server timestamps are little use for our analysis because they bunch
% together; it is common for several records to have identical server
% timestamps.
% However, server timestamps are useful for auditing because they should
% be relatively close in value to the client timestamps.
% For a small number of records (<100? \FILL{}), though, the client is off by a large
% margin: up to one week behind and up to one day ahead of the server.
% We attribute the large delays to issues when sending telemetry (lost internet
% connection, etc.) and smaller offsets to incorrect time settings on the client
% machine.

@@

Three important characteristics of the recorded data are the size of the codebases,
the length of the sessions, and the number of type
errors.
We discuss these in turn.
Codebase size and session length have long-tailed distributions
with many small items and few very large ones.
%% TODO check %% follow power-law distributions~\cite{csn-siamr-2009} 
%% https://aaronclauset.github.io/powerlaws/
%% bin version, is this relevant? sounds like no from abtract https://arxiv.org/pdf/1208.3524.pdf
The number of type analysis errors is somewhat low~(explained in the next
section), but sufficient to proceed with futher analysis.

@@

\begin{table}[t]\centering
  \caption{Size of analyzed code: number of files, number of lines, and lines in edit range}
  \label{t:codebase-size}
  % ;; (list* (max* vv) (median < vv) mm (stddev/mean mm vv) (percentile* vv)))
  % ;; percentile* = 0.95 -- 0.99
  % #hash((editrange . (1156036 926 3007043/817 30975.40553032358 (0.95 8220) (0.96 9858) (0.97 13656) (0.98 18956) (0.99 34725)))
  %       (event-count . (6079 138 38689/135 582.5752201972966 (0.95 960) (0.96 1174) (0.97 1304) (0.98 1836) (0.99 3302)))
  %       (files . (54884 7678 55257996/4639 11853.650098459446 (0.95 40029) (0.96 42771) (0.97 45417) (0.98 48688) (0.99 51761)))
  %       (lines . (1089963 3115 38928947/5992 22364.75846303099 (0.95 18561) (0.96 25892) (0.97 27014) (0.98 29725) (0.99 50547)))
  %       (timespan . (1387897376 845648 591343604248/185705 15724745.722577972 (0.95 10696064) (0.96 12715468) (0.97 15596940) (0.98 21153924) (0.99 35450460))))
  %% NOTE: 3 records have +1M lines, all from one nocheck session, all with 3 files
  %% - session "30,762,216,447,324"
  %% NOTE: 133 records have +52K files, all from one nocheck session
  %% - session "330,172,870,224,081" 
  %% (2nd biggest has 41K files "31,650,473,913,529")

@@

%% TODO figure --> table

\Cref{t:codebase-size} summarizes the size of projects in the dataset.
The top half of the figure reports the mean, standard deviation, median,
and \percentile{99} for the number of lines
and lines in edit range.\footnote{The data also contains the number of files in
each project, but these numbers are difficult to relate to typechecking because
projects may contain hundreds of files that define data assets.  For the
record, the median file count is 7,678 and the 99th percentile is 51,761 files.
}
The distributions zoom in zoomed-in distributions of the line and edit range
sizes.
For example, the x-axis of the first plot counts up to 1,000 lines and
the y-axis counts up to 1,000 telemetry records.

There is a huge amount of variation across projects.
The largest ones have over 50,000 lines of code
while the smallest have merely 1 line of code.
Unsurprisingly, the numbers come with large standard
deviations~([std] in~\cref{t:codebase-size}).
The median values are more reasonable, with roughly 3,000 lines of code
and 1,000 lines in edit ranges.

@@

\Cref{f:session-size} outlines the size of sessions using two metrics:
real time in seconds between the first and last record, and the total number of
records.
The top of~\cref{f:session-size} reports the mean, standard deviation,
median, and \percentile{99}; the bottom shows size distributions, zoomed in
to focus on at most 1,000 units (x-axis) and 1,000 sessions
(y-axis).\todo{Again, confused.}
